---
title: 'Twitter'
author: "Daniel Zeiner, Balint Szutor, Tao Yan, Mayssa Maring, Eudald Puig"
date: "04/10/2020"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

Introductory code:

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


Loading libraries:

```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```




# General Social Survey (GSS)

Loading data file:


```{r, read_gss_data, cache=TRUE}

setwd("C:/Users/mayss/Documents/LBS_Term1/Data Analytics for Finance/my_website/content/projects/project4")


gss <- read_csv((file="data/smallgss2016.csv"), 
                na = c("", "Don't know",
                       "No answer", "Not applicable"))

print(gss)
```

## Instagram and Snapchat, by sex

Creation of a  new variable, `snap_insta` that is *Yes* if the respondent reported using any of Snapchat (`snapchat`) or Instagram (`instagrm`), and *No* if not. 


```{r}

snap_insta <- gss %>% 
  mutate(
    snap_or_insta = ifelse( 
      snapchat=="Yes" | instagrm=="Yes", "Yes",
      ifelse(snapchat=="No" | instagrm=="No", "No", NA))
    
  )

print(snap_insta)
      
 

```



We calculate the proportion of Yes’s for `snap_insta` among those who answered the question, i.e. excluding NAs.


```{r}

yes_n <-snap_insta %>% 
 filter(snap_or_insta=="Yes") %>%  
  summarise(
    yes_count = n()
  ) 
  
 no_n <- snap_insta %>%
   
   filter(snap_or_insta=="No") %>% 
  summarise (
    no_count = n()
  )
  
proportion <- yes_n / (yes_n + no_n)
print(proportion)




```



We construct 95% CIs for men and women who used either Snapchat or Instagram

```{r}
get_CI_social <- function(n_sex){
  
new_data <- snap_insta %>% 
  filter(
    sex == n_sex,
    !is.na(snap_or_insta)
  ) %>% 
  select(snap_or_insta)

new_data

prop.test(count(new_data$snap_or_insta=="Yes"), 
          nrow(new_data),
          conf.level = 0.95)

p_hat <- count(new_data$snap_or_insta=="Yes") / nrow(new_data)

upper_CI <- p_hat + 1.96*sqrt((p_hat*(1-p_hat))/nrow(new_data))
names(upper_CI) <- "Upper CI"
lower_CI <- p_hat - 1.96*sqrt((p_hat*(1-p_hat))/nrow(new_data))
names(lower_CI) <- "Lower CI"

return(c(n_sex, upper_CI, lower_CI))
}

get_CI_social("Male")
get_CI_social("Female")

```



## Twitter, by education level

Can we estimate the *population* proportion of Twitter users by education level in 2016?. 

```{r}
gss %>% 
  filter(
    twitter == "Yes"
  ) %>% 
  group_by(
    degree
  ) %>% 
  summarise(
    proportion_degree = count(degree)
  ) %>% 
  mutate(
    proportion_degree = proportion_degree / sum(proportion_degree)
  )
```

We Turn `degree` from a character variable into a factor variable:

```{r}
gss <- gss %>% 
  mutate(
    degree = factor(degree, levels=c("Lt high school", "High school", "Junior college", "Bachelor", "Graduate", "NA"))
  )
```


We create a  new variable, `bachelor_graduate` that is *Yes* if the respondent has either a `Bachelor` or `Graduate` degree.

```{r}
bachelor_graduate_data <- gss %>% 
  mutate(
    bachelor_graduate = ifelse(degree == "Bachelor" | degree == "Graduate" & !is.na(degree), "Yes", "No")
  )
```


We calculate the proportion of `bachelor_graduate` who do (Yes) and who don't (No) use twitter. 

```{r}
bachelor_graduate_data %>% 
  filter(twitter=="Yes") %>% 
  select(bachelor_graduate) %>% 
  group_by(bachelor_graduate) %>% 
  summarise(
    proportion_bach = count(bachelor_graduate)
  ) %>% 
  mutate(
    proportion_bach = proportion_bach / sum(proportion_bach)
  )
```

We construct two 95% CIs for `bachelor_graduate` vs whether they use (Yes) and don't (No) use twitter. 

```{r}
bach_grad_twitter <- bachelor_graduate_data %>% 
  filter(
    twitter == "Yes",
    !is.na(bachelor_graduate)
  ) %>% 
  select(bachelor_graduate)

p_hat <- count(bach_grad_twitter$bachelor_graduate=="Yes") / nrow(bach_grad_twitter)

upper_CI <- p_hat + 1.96*sqrt((p_hat*(1-p_hat))/nrow(bach_grad_twitter))
names(upper_CI) <- "Upper CI"
lower_CI <- p_hat - 1.96*sqrt((p_hat*(1-p_hat))/nrow(bach_grad_twitter))
names(lower_CI) <- "Lower CI"

upper_CI
lower_CI

bach_grad_twitter <- bachelor_graduate_data %>% 
  filter(
    twitter == "No",
    !is.na(bachelor_graduate)
  ) %>% 
  select(bachelor_graduate)

p_hat <- count(bach_grad_twitter$bachelor_graduate=="Yes") / nrow(bach_grad_twitter)

upper_CI <- p_hat + 1.96*sqrt((p_hat*(1-p_hat))/nrow(bach_grad_twitter))
names(upper_CI) <- "Upper CI"
lower_CI <- p_hat - 1.96*sqrt((p_hat*(1-p_hat))/nrow(bach_grad_twitter))
names(lower_CI) <- "Lower CI"

upper_CI
lower_CI
```


Do these two Confidence Intervals overlap?

The two intervals do not overlap, which could be because we have large enough sample size that we are confident in our observations.

